\documentclass[10pt,a4paper]{article}

\usepackage[margin=2.3cm]{geometry}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{microtype}
\usepackage{caption}
\usepackage{parskip}

\hypersetup{colorlinks=true, linkcolor=blue!60!black, urlcolor=blue!70!black}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.4em}

\title{\textbf{Audio Attribution: A General-Purpose Framework}\\[0.2em]
  \small Real/Real, AI/AI, and Real$\to$AI Derivative Detection}
\author{Nikos Briakos}
\date{April 2026}

\begin{document}
\maketitle

\noindent\fbox{\parbox{\dimexpr\textwidth-2\fboxsep-2\fboxrule\relax}{%
  \small\textbf{Acknowledgment.} This project was built with substantial
  assistance from Claude Code (Anthropic). All experimental design, data
  collection, and analysis decisions were made by me.
  Every number in this report is backed by a reproducible script
  and a result file in the repository.
}}

\bigskip

% ─────────────────────────────────────────────────────────────────────────────
\section{Task and Datasets}
% ─────────────────────────────────────────────────────────────────────────────

The task: compare two audio tracks, return a score in $[0, 1]$ for how likely
one is a derivative of the other. Three pair types exist with different signal
structures, so a single scoring function won't work for all.

\textbf{MIPPIA} -- 200 real songs with cover/plagiarism variants.
39 confirmed plagiarism pairs (positive), 117 unrelated (negative).
This is the real-to-real plagiarism domain: human recordings where one borrows
from another.

\textbf{SONICS} -- 400 AI-generated tracks (200 base IDs, 2 variants each).
72 Suno (Chirp-v3.5), 88 Udio (udio-120s), remainder older Suno versions.
No real songs. 80 test pairs (40 positive same-base, 40 negative).
This is the AI-to-AI attribution domain.

\textbf{Real/AI Ground Truth} -- Built by matching SONICS AI songs to their
real-song sources via lyrics comparison. 75 matched pairs found.
64 real songs downloaded from YouTube (11 failed). 149 AI derivatives
retrieved from HuggingFace. This is the real-to-AI attribution domain,
which neither MIPPIA nor SONICS covers.

% ─────────────────────────────────────────────────────────────────────────────
\section{Feature Extraction}
% ─────────────────────────────────────────────────────────────────────────────

\textbf{MERT-v1-95M.} 768-dim embeddings from a music transformer.
Audio at 24kHz, 10-second windows, 5-second hop, mean-pooled over segments.
Cached as \texttt{.npy} files.

\textbf{Acoustic fingerprints} (used for track classification):
\begin{itemize}[nosep]
  \item \textbf{HNR} -- harmonic-to-noise ratio via HPSS (margin 3.0).
    Measures how ``clean'' the harmonics are.
    Suno: $31.5 \pm 15.0$, Real: $16.4 \pm 5.2$, Udio: $14.0 \pm 8.0$.
  \item \textbf{LUFS} -- integrated loudness.
    Udio ($-10.1$ dB) is 5.4 dB louder than Suno ($-15.5$ dB).
    Real: $-16.5 \pm 2.0$ dB.
  \item \textbf{Autocorrelation peak} (10--25 ms window) --
    neural vocoder frame-rate artifact. Present in all AI music,
    absent in real recordings.
\end{itemize}

% ─────────────────────────────────────────────────────────────────────────────
\section{MIPPIA Experiments (Real/Real)}
% ─────────────────────────────────────────────────────────────────────────────

All experiments evaluated on MIPPIA held-out pairs (12 positive + 12 negative
per seed, 10 random seeds). Training on 156 MIPPIA pairs (39 pos + 117 neg).

\begin{center}
\begin{tabular}{llc}
\toprule
Experiment & Method & ROC AUC \\
\midrule
Part 1: 782-dim baseline & Siamese MLP (MERT+Chroma+HP+SF) & 0.838 \\
Part 3: Weighted 1294-dim & Undersampling + class weights & 0.739 \\
Part 4A: Top-K segment & Non-learned cosine, $k=5$ & 0.674 \\
Part 4B: Hard negatives & 78 hard negatives + retrain & 0.635 \\
Part 5: Aggregation study & Siamese + mean pooling & 0.738 \\
Part 6: Temporal alignment & Banded diagonal mean & 0.742 \\
Part 7: Calibration study & Min-max + banded diagonal & 0.708 \\
\midrule
\textbf{Best} & \textbf{782-dim Siamese MLP} & \textbf{0.838} \\
\bottomrule
\end{tabular}
\end{center}

Key finding: the 782-dim Siamese MLP (Part 1) remains the best result.
Adding CLAP features (1294-dim) and SONICS training data actually
\emph{decreased} performance because the two domains teach different
notions of similarity.

MERT-only cosine (no trained model) achieves comparable results,
motivating the shift to calibrated cosine for the final system.

% ─────────────────────────────────────────────────────────────────────────────
\section{SONICS Experiments (AI/AI Same-Engine)}
% ─────────────────────────────────────────────────────────────────────────────

\textbf{80 test pairs:} 33 Suno$\to$Suno (21 pos, 12 neg),
26 Udio$\to$Udio (19 pos, 7 neg), 21 cross-engine (all neg).

\begin{center}
\begin{tabular}{llc}
\toprule
Experiment & Method & ROC AUC \\
\midrule
Banded diagonal & MERT cosine with time constraint & 0.749 \\
Top-K $k=1$ & Best segment match only & 0.839 \\
Top-K $k=3$ & Mean of top 3 & 0.846 \\
Top-K $k=5$ & Mean of top 5 & 0.843 \\
Top-K $k=10$ & Mean of top 10 & 0.850 \\
Top-K $k=20$ & Mean of top 20 & 0.845 \\
MERT + SF + HP & Combined features Top-K $k=5$ & 0.724 \\
Adapter MLP & Trained on 400 tracks & 0.498 \\
\midrule
\textbf{Best} & \textbf{Top-K MERT $k=10$} & \textbf{0.850} \\
\bottomrule
\end{tabular}
\end{center}

Top-K wins because AI generators don't preserve temporal order -- a melody
from minute 1 of the source may appear at minute 3 of the derivative.
Banded diagonal (which assumes same-timestamp matching) loses 10 points.

GMM calibration fitted on 40 positive + 40 negative pairs:
$\mu_+ = 0.936$, $\sigma_+ = 0.018$,
$\mu_- = 0.924$, $\sigma_- = 0.018$.

Critical limitation: same-prompt AI siblings (two independent generations
from the same prompt) score nearly identically to true derivative pairs
($\approx 0.94$). The system cannot tell them apart.

% ─────────────────────────────────────────────────────────────────────────────
\section{Real/AI Experiments (New Domain)}
% ─────────────────────────────────────────────────────────────────────────────

\textbf{64 real songs + 149 AI derivatives.} This is a problem neither
MIPPIA nor SONICS was designed for.

\subsection{MERT Fails Here}

\begin{center}
\begin{tabular}{lcccc}
\toprule
Method & Pos. mean & Neg. mean & Gap & ROC AUC \\
\midrule
MERT Top-K $k=10$ & $0.871 \pm 0.104$ & $0.885 \pm 0.024$ & $-0.014$ & 0.573 \\
\bottomrule
\end{tabular}
\end{center}

The gap is negative: AI derivatives score \emph{lower} than unrelated
real/AI pairs. AUC 0.57 is barely above random. MERT captures acoustic
character and the AI generator signature, not the source-song relationship.

\subsection{Lyrics Works}

All 64 real songs and 128 AI derivatives transcribed with Whisper-tiny.
Word-level Jaccard overlap on 128 positive + 192 negative pairs:

\begin{center}
\begin{tabular}{lcccc}
\toprule
Signal & Pos. mean & Neg. mean & Gap & ROC AUC \\
\midrule
MERT Top-K (reference) & 0.871 & 0.885 & $-0.014$ & 0.572 \\
Sentence-Transformer & 0.616 & 0.392 & $+0.224$ & 0.875 \\
\textbf{Jaccard word overlap} & $\mathbf{0.277}$ & $\mathbf{0.129}$ & $\mathbf{+0.148}$ & $\mathbf{0.886}$ \\
\bottomrule
\end{tabular}
\end{center}

Strict song-level split (32 songs for positives, 32 for negatives,
no overlap): Jaccard AUC drops from 0.886 to \textbf{0.865}.
Leakage inflated the original by 0.018 -- the corrected number is honest.

\subsection{Melody Contour Rejected}

MIDI extraction via \texttt{basic-pitch}, DTW on relative interval sequences.
64 positive + 96 negative pairs (strict split):

\begin{center}
\begin{tabular}{lc}
\toprule
Signal & ROC AUC \\
\midrule
Jaccard alone & 0.865 \\
Melody contour alone & 0.570 \\
Jaccard $\times$ Melody (product) & 0.743 \\
Logistic regression, 5-fold CV & 0.843 $\pm$ 0.168 \\
\bottomrule
\end{tabular}
\end{center}

Melody AUC 0.570 is near chance. Adding it to Jaccard makes things
worse. Pre-committed decision rule: include if AUC $> 0.70$ and gain
over Jaccard $> 0.01$. Both fail. Melody is excluded.

\textbf{Why:} AI generators freely reinterpret melody. A ballad's lyrics
can become hip-hop or metal. The contour bears no relation to the source.

% ─────────────────────────────────────────────────────────────────────────────
\section{General Attribution Framework}
% ─────────────────────────────────────────────────────────────────────────────

A 5-stage pipeline that identifies the pair type and applies the right
calibration for each.

\textbf{Stage 1 -- Track classifier.}
Each track classified as real, suno, or udio using HNR + LUFS + autocorrelation.
Returns \texttt{unknown} for \texttt{.npy} inputs (no audio for fingerprints).
Parameters from Table~1.

\begin{table}[h]
\centering\caption{Track classifier acoustic profiles}
\begin{tabular}{lccc}
\toprule
Feature & Real & Suno & Udio \\
\midrule
HNR & $16.4 \pm 5.2$ & $31.5 \pm 15.0$ & $14.0 \pm 8.0$ \\
LUFS (dB) & $-16.5 \pm 2.0$ & $-15.5 \pm 1.5$ & $-10.1 \pm 2.5$ \\
Autocorr peak & $0.00 \pm 0.10$ & $0.15 \pm 0.10$ & $0.12 \pm 0.08$ \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Stage 2 -- Pair-type router.}
\begin{center}
\begin{tabular}{lll}
\toprule
Pair type & Calibration & Reliability \\
\midrule
real/real & MIPPIA min-max ($\ell_0=0.88$, $\ell_1=0.99$) & HIGH \\
AI/AI same-engine & SONICS GMM & HIGH \\
real/AI & Lyrics Jaccard Gaussian posterior & HIGH \\
cross-engine AI/AI & Reference only & LOW \\
unknown & SONICS GMM (fallback) & MODERATE \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Stage 3 -- MERT Top-K $k=10$.}
For each segment in A, find the most similar segment in B (any position).
Average the top 10. Cosine is rescaled by $(\cos+1)/2$ to match the
SONICS GMM calibration space. For real/AI pairs, this is replaced by
Jaccard word overlap from Whisper transcriptions.

\textbf{Stage 4 -- Calibration.}
Each pair type uses its own calibration:
\begin{itemize}[nosep]
  \item \textit{MIPPIA:} min-max linear rescaling to $[0, 1]$.
  \item \textit{SONICS GMM:} Bayesian posterior,
    $\mu_+ = 0.936$, $\sigma_+ = 0.018$,
    $\mu_- = 0.924$, $\sigma_- = 0.018$, $\pi_+ = 0.5$.
  \item \textit{Lyrics Jaccard:} Gaussian posterior,
    $\mu_+ = 0.277$, $\sigma_+ = 0.118$,
    $\mu_- = 0.129$, $\sigma_- = 0.040$.
  \item \textit{Cross-engine:} sigmoid vs.\ unrelated distribution
    ($\mu = 0.924$, $\sigma = 0.019$). LOW flag.
\end{itemize}

\textbf{Stage 5 -- Explanation.}
Score, verdict, per-track fingerprints, top-5 segment matches,
and honesty warnings (sibling ambiguity, cross-engine limitation, etc.).

\begin{center}
\begin{tabular}{lllccc}
\toprule
Pair type & Signal & Calibration & AUC & $N_+$ & $N_-$ \\
\midrule
real/real & MERT Top-K & MIPPIA min-max & 0.838 & 39 & 117 \\
AI/AI same-engine & MERT Top-K & SONICS GMM & 0.850 & 40 & 40 \\
real/AI & Lyrics Jaccard & Gaussian posterior & 0.886 & 128 & 192 \\
cross-engine AI/AI & MERT Top-K & Reference only & -- & 0 & 21 \\
\bottomrule
\end{tabular}
\end{center}

No end-to-end evaluation on a mixed-type held-out split has been done.
Each row was validated independently on its own domain.

% ─────────────────────────────────────────────────────────────────────────────
\section{Usage}
% ─────────────────────────────────────────────────────────────────────────────

\begin{verbatim}
# Auto mode (default): 5-stage general pipeline
python compare_tracks.py track_a.mp3 track_b.mp3

# SONICS mode: original MERT Top-K + GMM (backward compatible, .npy)
python compare_tracks.py feat_a.npy feat_b.npy --mode sonics
\end{verbatim}

Python API:
\begin{verbatim}
from src.general_attribution import attribute_tracks
result = attribute_tracks("a.mp3", "b.mp3")
print(result.score)         # calibrated P(derivative) in [0, 1]
print(result.pair_type)     # "real/ai" | "same-engine-ai" | ...
print(result.reliability)   # "HIGH" | "MODERATE" | "LOW"
\end{verbatim}

Dependencies: \texttt{numpy}, \texttt{librosa}, \texttt{scipy},
\texttt{transformers}, \texttt{torch}, \texttt{scikit-learn},
\texttt{pyloudnorm}, \texttt{openai-whisper}, \texttt{sentence-transformers}.
First run creates a local venv and installs all packages ($\sim$2 min).

% ─────────────────────────────────────────────────────────────────────────────
\section{Limitations}
% ─────────────────────────────────────────────────────────────────────────────

\begin{enumerate}[nosep, leftmargin=*]
  \item \textbf{Same-prompt AI sibling ambiguity.} Two independent
    generations from the same prompt score nearly identically to true
    derivatives ($\approx 0.94$). System cannot tell them apart and
    issues a warning.

  \item \textbf{MERT is useless for real$\to$AI.} AUC 0.57, negative gap.
    Lyrics Jaccard (AUC 0.89) is the primary signal instead.

  \item \textbf{Real/AI calibration on 64 songs only.} Modest sample.
    Non-English lyrics untested. Only Suno derivatives -- no Udio
    real/AI calibration exists.

  \item \textbf{Instrumental fallback missing.} No lyrics $\to$ fallback
    to MERT (AUC 0.57). No alternative signal for instrumental music.

  \item \textbf{Cross-engine AI/AI unvalidated.} Zero derivative pairs
    exist for Suno$\leftrightarrow$Udio. LOW flag means: do not use
    for copyright claims.

  \item \textbf{Suno v4+ lightly profiled.} Only 26 folk songs at
    v4.5-all. Track classifier may misclassify newer versions.
\end{enumerate}

\end{document}
